---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# iteval

<!-- badges: start -->
<!-- badges: end -->

The goal of iteval is to ease implementation of the discrimination and
calibration measures for predicted individualized treatment effect as
described by Hoogland et al. DOI: [10.1002/sim.10186](https://doi.org/10.1002/sim.10186). 

#### Contributions

Many thanks to [Kim Daniel Jakobsen](https://github.com/kjakobse) for writing a fast C++ implementation of mbcb that allows for sampling weights and provides asymptotic se's (based on section 5 in Quade, D. (1971). Nonparametric partial correlation. Stichting Mathematisch Centrum. Mathematical Statistics. Stichting Mathematisch Centrum; implemented analogous to the cidxcn subroutine from the Hmisc R package (https://github.com/harrelfe/Hmisc/blob/master/src/ratfor/cidxcn.r).

## Installation

You can install the development version of **iteval** from [GitHub](https://github.com/) with:

``` {r eval=F}
# install.packages("devtools")
devtools::install_github("jeroenhoogland/iteval")
```

and load the package using 

```{r}
library(iteval)
```

## Example

This is a basic example which illustrates usage of the main functions. For this 
purpose, we first simulate some data from a data generating mechanism with a 
single covariate `x`, a treatment indicator `trt`, an interaction between the two
and observed binary outcome `y` and fit a basic logistic regression model.

```{r example}
# generate some  data
set.seed(123) # for replicability
n <- 500
x <- rnorm(n) # covariate
trt <- rbinom(n, 1, .5) # treatment
beta <- c(0, 1, 1, -.75) # coefficients
lp <- cbind(1, x, trt, x*trt) %*% beta # linear predictor
y <- rbinom(n, 1, plogis(lp)) # simulated y

# model
mod <- glm(y ~ x * trt, family = "binomial")
```

Next, we obtain predicted of y under control treatment (`y0hat`), treated condition
(`y1hat`) and predicted ite's (`deltahat`).

```{r}
# Predicited ite's
control.data <- cbind(1, x, 0, 0) 
treated.data <- cbind(1, x, 1, x)

# prediction under the control condition
y0hat <- plogis(control.data %*% coef(mod))

# prediction under the treated condition
y1hat <- plogis(treated.data %*% coef(mod))

# and their difference
deltahat <- y1hat - y0hat

# true outcome probabilities given data generating mechanism
p0 <- plogis(control.data %*% c(0, 1, 1, -.75))
p1 <- plogis(treated.data %*% c(0, 1, 1, -.75))
```

Subsequently, we apply the performance measures in the simulated data. NB.: note that these
estimates will be optimistic since this same data set was used to fit the prediction 
model; in practice, indepdenet data or an internal validation procedure is required.

### Measures of discriminative performance

Each of the measures need to be aware of the treatment assignments. `ind.A` contains
the indices for the controls and `ind.B` contains the indices for the treated.

```{r}
ind.A <- which(trt == 0)
ind.B <- which(trt == 1)
```

All is in place now. The treatment assignments are as follows:

```{r}
table(trt)
```

Hence, 1:1 matching excluded some individuals and repeated subsampling of the larger
group is used more of the data (`nresample=100`).

```{r}
# cbendelta
res <- cbendelta(deltahat, y, ind.A, ind.B, nresample = 250, get.all=TRUE)
mean(res) # cbendelta
hist(res, xlab="cbendelta", main=""); abline(v=mean(res), col="red")

# cbeny0
res <- cbeny0(y0hat, y1hat, y, ind.A, ind.B, "y0hat", nresample = 250, get.all=TRUE)
mean(res) # cbendelta
hist(res, xlab="cbeny0", main=""); abline(v=mean(res), col="red")

# mbcb (does not depend on matching)
mbcb(y0hat, y1hat)

# 'true' mbcb for this model (possible since sim setting)
mbcb(y0hat, y1hat, y0hat.updated = p0, y1hat.updated = p1)
```


### Calibration performance

```{r}
# Example 'true' calibration (possible since sim setting)
# otherwise, cal() takes estimates p0 and p1 based on 
# independent data.
(cal <- cal(y0hat,y1hat,y,trt,p0,p1,type="both"))
```

